Nature or Nurture? Collisionless Evolution of Galactic Disc-Halo Systems

### 2019-09-23 Thesis/Introduction.tex
Final commit

A model of the Universe one might consider is one comprised of cold dark matter and dark energy, described on large scales by a single parameter, $\Lambda$. This paradigm is called $\Lambda$CDM and is the prevailing model in the astronomy community. The model has been enormously successful in explaining observations, including the CMB power spectrum. While there remain open problems with $\Lambda$CDM, it appears to predict many properties on Galactic and extragalactic scales correctly \citep{kolb_turner,dodelson, BT}. Proposals to modify $\Lambda$CDM tend to present only slight variations on the model, such as allowing for self interacting dark matter (SIDM).
\begin{figure}
	\includegraphics[width=\textwidth]{../figures/local_group}
	\caption{A map of the Local Group obtained per the license in \citet{local_group_map}.}\label{fig:local_group}
\end{figure}

@@

The baryons, interacting through baryon acoustic oscillations (BAO), are delayed in falling into the potential wells made by the dark matter. As the baryons fall into the centers of the potential wells over time, the cores of the dark matter distributions, comprised of spheroidal structures called halos, contract. After infall into a dark matter halo, the baryons cool to the point where stars are able to form. This is how the first galaxies formed.

Over time, galaxies begin to shift away from the linear mapping between perturbation and galaxy. Their forces on each other start to matter locally, and galaxies begin to merge hierarchically. Collections of galaxies form, sometimes massive clusters of $10^{15}$ solar masses. Other times, groups with a couple massive galaxies become host to many smaller galaxies called subhalos. Other galaxies belong to no group or cluster, and exist in the expansive void between clusters. Dark energy continues to pull the Universe apart, while locally galaxies are kept together by dark matter. Merging of galaxies continues on local scales as larger galaxies become larger, bringing us to today.

We live in a spiral galaxy of approximately $10^{12}$ solar masses (dark matter + baryons), in a group with another galaxy, M31 (Andromeda), which has a roughly similar mass. Our Local Group (Fig. \ref{fig:local_group}), as it is called, is comprised of many smaller galaxies. In our Milky Way's halo, for instance, a merger is beginning between the Milky Way and the Large and Small Magellanic Clouds (LMC/SMC), shown in the third quadrant of the inner circle around the Milky Way in Fig. \ref{fig:local_group}. The SMC/LMC systems total approximately 10-20\% of the Milky Way's mass in total \citep{erkal_lmc}. Most of our subhalos are not nearly this massive, although they number in the hundreds. The picture painted of the Local Group is that of M31 and the Milky Way devouring smaller galaxies. Of course, this will have a substantial impact on the evolution of the stars in the Milky Way, and is the crux of this thesis. By studying the Milky Way's local environment, we can hope to understand the interactions dark matter has on baryons, and perhaps infer something about how dark matter is distributed in our Galaxy.

@@

Many of the hardest questions in Galactic astrophysics stem from the fact that dark matter is not directly observable. As an example, what is the underlying smooth distribution of dark mass? How many dark matter subhalos should the Milky Way have? What is the effect on large scales if we choose something other than $\Lambda$CDM? While these questions are not easily answerable by direct observation, we can complement indirect observations with a solid theoretical understanding to reach confident answers.

@@

Over the last four decades, many tools have been developed to study the structure of the present day Universe. These tools have become increasingly sophisticated as computation power and algorithms research advance. One tool is an \textit{ab initio} cosmological simulation, which simulates the entire formation history of the Universe from some informed initial density field.  This initial density field is derived from the primordial power spectrum and early-Universe baryon physics, and a sample can be drawn and evolved with periodic boundary conditions \citep{music}. If done correctly, this creates a representative sample of many structures in a $\Lambda$CDM Universe. Halos evolving in these simulations feel the full effect of the completely simulated cosmological environment. Such simulations stand in contrast to isolated galaxy simulations, which seek to explain the detailed dynamics arising from generic effects in individual galaxies. These simulations start with a prescription for what an equilibrium galaxy looks like, and study the evolution of galaxies initialized from these prescriptions. While missing some critical cosmological physics, these simulations may be run at very high resolution.

@@

We broadly commented on the hierarchical formation of galaxies. Over time, galaxies that merge with other galaxies should become well-mixed in the host halo. That is, there should be an underlying smooth distribution of mass in which all of the subhalos reside. A key prediction of $\Lambda$CDM is that this smooth distribution should have a universal density profile in the limit that dark matter dominates the Universe \citep{nfw}. Furthermore, the dark matter should have a density distribution somewhat resembling a squashed football, a spheroid flattened on two axes, in general. There are, in fact, the special cases of being either perfectly spherical, prolate, or oblate. Whether or not this picture holds observationally in real galaxies is an open problem. The density profile proposed by \citet{nfw} is cuspy at the center, whereas some observations are more consistent with a flatter central density \citep{de_blok_core_cusp}. Astronomers have dubbed this the Core-Cusp Problem.

Besides the Core-Cusp Problem in the inner parts of dark matter haloes, one thing that is clear is that $\Lambda$CDM halos should have many subhalos. As the general theory for the formation of galaxies goes, these subhalos were at one point independent perturbations that have since merged into the Milky Way's halo, and should have their own stellar and gas content. This means that we expect many massive subhalos to be directly observable. Another triumph of simulations is putting an actual number on how many subhalos we expect from $\Lambda$CDM. Surprisingly, far fewer subhalos are actually observed \citep{mooresubhalos, Klypin1999, springel2008}. This discrepancy has since been dubbed the Missing Satellite Problem, and is an open problem for $\Lambda$CDM and astronomy.

@@

There is also a stellar component of the halo, which contains remnants of old mergers and star clusters known as globular clusters. This section is organized by talking about tidal disruption of subhalos in the Galactic halo, the dynamics of the central part of the Galaxy, and vertical structure in the Galactic disc. How the cosmological environment effects these observables is the bulk of this thesis, and we motivate here why these aspects are the subject of focus.

@@

In the hierarchical formation of galaxies, smaller galaxies merge with larger galaxies and become a part of the larger galaxy's halo. As subhalos become more a part of the host system, they experience tidal forces that stretch and squeeze their mass content. Dark matter and stellar material is stripped from these merging galaxies over time to form streams of tidal debris. In the long run, this debris will equlibrate with the stellar halo. It should be noted that the same effects apply to globular clusters too.

@@

The orbits of streams are not the only way to learn about the Milky Way's dark matter halo. It is worth mentioning that there is the possibility of using the fine structure of stellar streams to learn more about the subhalo population in the halo.  Gaps form when subhalos pass through tidal streams, and this is one direction that is being explored for leveraging stellar stream data (e.g. \citet{erkal_2016_stream_gaps}).

@@

The Milky Way's massive subhalos are most likely inducing non-planar density patterns in the disc, and we would like to understand the nature of these patterns through simulation. Unfortunately, we need the high resolution of isolated galaxy simulations to study these effects, and the realism of an \textit{ab initio} Galactic halo.  This has motivated a long history of attempts to bridge the gap between {\it ab initio} cosmological simulations and those of isolated galaxies.  A particularly simple class of simulations involves the interaction of a disc
with a single satellite galaxy or dark matter subhalo.  For example \citet{kazantzidis2008} perform simulations in which a thin Milky Way-like disc is subjected to a series of encounters with a satellite.  The masses and orbital parameters of the satellites are motivated by substructure found in a halo appropriate to a Milky Way-like galaxy from a cosmological simulation.  Their simulation demonstrated that satellite encounters lead to general disc heating as well as distinctive disc features such as bars, spiral structure, flares, and rings.  \citet{purcell2011} model the response of the Milky Way to the gravitational effects of
the Sgr dSph by simulating disc-dwarf encounter for different choices of the Sgr progenitor's mass.  Their conclusion is that Sgr may have triggered the development the spiral structure seen today in the Milky Way.  This approach has gained considerable traction in the last couple years, with authors using single-satellite encounters to explain the existence of low-latitude streams and other vertical structure in the Milky Way's disc \citep[for example]{widrow_2014, dlv_2015, donghia_2016, laporte_2016, laporte_2018}. One particularly interesting feature arising from these works is that stars may occasionally be kicked out of the galactic disc to form kicked-up disc populations \citep{johnston_kud_review, laporte_2019_feathers}. Such populations may explain low-latitude Milky Way structures, and possibly kinematic data of M31 \citep{dorman_2013_m31}.

@@

Of course, the disc of the Milky Way lives within a population of satellite galaxies and, quite possibly, pure dark matter subhalos, and single-satellite encounters do not describe the complexity of the cosmological environment \citep{Klypin1999,mooresubhalos,springel2008}.  With this in mind \citet{Font2001} simulated the evolution of an isolated disc-bulge-halo model where the halo was populated by several hundred subhalos.  They concluded that that substructure played only a minor role in heating the disc, a result that would seem in conflict with those of the \citet{kazantzidis2008} that would come later.  Numerical simulations by \citet{gauthier2006, dubinski2008} sheds some light on this discrepancy.  In these simulations, 10\% of the the halo mass in an isolated disc-bulge-halo system is replaced by subhalos with a mass distribution motivated by the cosmological studies of \citet{gao2004}.  In the \citet{gauthier2006} simulation, little disc heating occurs during the first 5 Gyr, at which point satellite interactions provoke the formation of a strong bar, which in turn leads to significant heating.  Not surprisingly, when the experiment is repeated with different initial conditions for the satellites the timing of bar formation can vary.  These experiments also suggest that a large fraction of the mass initially in subhalos is tidally stripped leaving behind a complex system of tidal debris.

The aforementioned simulations, although more realistic than single-satellite encounters, have three main problems.  First, they don't allow for halo triaxiality.  Secondly, the disc is initialized at its present-day mass whereas real stellar discs form over several Gyr.  Finally, the subhalo populations are inserted into the halo by hand.  While the mass and spatial distributions of the subhalos are motivated by cosmological simulations, they may not capture important properties of realistic halos. This has historically been solved by modifying a traditional \textit{ab initio} simulation. The zoom-in technique (a Monte Carlo adaptive mesh) is used to study the evolution of a disc in a single halo at high resolution.  This is accomplished by resampling the initial conditions at a sequence of higher resolutions after identifying an area of interest. With sufficiently realistic feedback, impressively realistic results can be obtained. For instance, \citet{gomez_2017} presented a study of vertical structure in such simulations. In these simulations, the authors were able to look at several Milky Way-like galaxies in a variety of environments at resolutions that rival isolated galaxy simulations. These realistic studies have opened the door to a wide variety of questions pertaining to the Milky Way's interaction with the cosmological environment.

@@

The general idea for growing a disc is to divide a zoom-in simulation into three phases.  During the first, a cosmological simulation of a region of interest is run, and a halo is identified for study.  A rigid disc potential is then slowly grown in this halo, allowing the halo particles to respond adiabatically to its time-varying potential.  Finally, the rigid disc is replaced by live particles and the final phase of the simulation proceeds with live disc and halo particles.  \citet{debuhr_2012} use this scheme to introduce stellar discs into dark matter halos from the Aquarius Project \citep{springel2008}.  They add a rigid disc potential at redshift $z=1.3$ with a mass parameter for the disc that grows linearly with the scale factor until $z=1$ when it reaches its final value.  The disc potential is initially centered on the potential minimum of the halo and oriented so that its symmetry axis points along either the minor or major axis of the halo.  During the rigid disc phase, the centre of the potential moves due to its interaction with the halo (Newton's 3rd law) so that linear momentum is conserved.  However, the orientation of the disc is held fixed and therefore angular momentum isn't.  To initialize the live disc, \citet{debuhr_2012} approximate the halo potential as a flattened logarithmic potential and then determine the disc distribution function by solving the Jeans equations in axisymmetry.

@@

The thesis is organized in the following manner. In Chapter~\ref{ch:background}, we reduce the necessary theoretical background to understand this work to a succinct summary. Following this, in Chapter~\ref{ch:paper_i}, we explore the first peer-reviewed paper on which this thesis is based. It outlines the disc insertion scheme and comments on how the dynamic nature of the Milky Way's stellar disc should affect its inner halo. Chapter~\ref{ch:paper_ii} presents the second published paper which focuses on the question of how stable disks are against forming bars in a cosmological setting. A final paper is presented in Chapter~\ref{ch:paper_iii}, where we look at how disks form vertical structure in a cosmological setting. We conclude with summaries and ideas for future work in Chapter~\ref{ch:conclusion}.

>>>>

A model of the Universe one might consider is one comprised of cold dark matter and dark energy, described on large scales by a single parameter, $\Lambda$. This paradigm is called $\Lambda$CDM and is the prevailing model in the astronomy community. The model has been enormously successful in explaining observations, including the CMB power spectrum. While there remain open problems with $\Lambda$CDM, it appears to predict many properties on Galactic and extragalactic scales correctly \citep{kolb_turner,dodelson, BT}. Proposals to modify $\Lambda$CDM tend to present only slight variations on the model, such as allowing for self interacting dark matter (SIDM).
\begin{figure}
	\includegraphics[width=\textwidth]{../figures/local_group}
	\caption{A map of the Local Group obtained from \citet{local_group_map}, and reproduced under the Creative Commons Attribution-Share Alike 2.5 Generic license.}\label{fig:local_group}
\end{figure}

@@

The baryons, interacting through baryon acoustic oscillations (BAO), are delayed in falling into the potential wells made by the dark matter. As the baryons fall into the centers of the potential wells over time, the cores of the dark matter distributions, comprised of spheroidal structures called haloes, contract. After infall into a dark matter halo, the baryons cool to the point where stars are able to form. This is how the first galaxies formed.

Over time, galaxies begin to shift away from the linear mapping between perturbation and galaxy. Their forces on each other start to matter locally, and galaxies begin to merge hierarchically. Collections of galaxies form, sometimes massive clusters of $10^{15}$ solar masses. Other times, groups with a couple massive galaxies become host to many smaller galaxies called subhaloes. Other galaxies belong to no group or cluster, and exist in the expansive void between clusters. Dark energy continues to pull the Universe apart, while locally galaxies are kept together by dark matter. Merging of galaxies continues on local scales as larger galaxies become larger, bringing us to today.

We live in a spiral galaxy of approximately $10^{12}$ solar masses (dark matter + baryons), in a group with another galaxy, M31 (Andromeda), which has a roughly similar mass. Our Local Group (Fig. \ref{fig:local_group}), as it is called, is comprised of many smaller galaxies. In our Milky Way's halo, for instance, a merger is beginning between the Milky Way and the Large and Small Magellanic Clouds (LMC/SMC), shown in the third quadrant of the inner circle around the Milky Way in Fig. \ref{fig:local_group}. The SMC/LMC systems total approximately 10-20\% of the Milky Way's mass in total \citep{erkal_lmc}. Most of our subhaloes are not nearly this massive, although they number in the hundreds. The picture painted of the Local Group is that of M31 and the Milky Way devouring smaller galaxies. Of course, this will have a substantial impact on the evolution of the stars in the Milky Way, and is the crux of this thesis. By studying the Milky Way's local environment, we can hope to understand the interactions dark matter has on baryons, and perhaps infer something about how dark matter is distributed in our Galaxy.

@@

Many of the hardest questions in Galactic astrophysics stem from the fact that dark matter is not directly observable. As an example, what is the underlying smooth distribution of dark mass? How many dark matter subhaloes should the Milky Way have? What is the effect on large scales if we choose something other than $\Lambda$CDM? While these questions are not easily answerable by direct observation, we can complement indirect observations with a solid theoretical understanding to reach confident answers.

@@

Over the last four decades, many tools have been developed to study the structure of the present day Universe. These tools have become increasingly sophisticated as computation power and algorithms research advance. One tool is an \textit{ab initio} cosmological simulation, which simulates the entire formation history of the Universe from some informed initial density field.  This initial density field is derived from the primordial power spectrum and early-Universe baryon physics, and a sample can be drawn and evolved with periodic boundary conditions \citep{music}. If done correctly, this creates a representative sample of many structures in a $\Lambda$CDM Universe. Haloes evolving in these simulations feel the full effect of the completely simulated cosmological environment. Such simulations stand in contrast to isolated galaxy simulations, which seek to explain the detailed dynamics arising from generic effects in individual galaxies. These simulations start with a prescription for what an equilibrium galaxy looks like, and study the evolution of galaxies initialized from these prescriptions. While missing some critical cosmological physics, these simulations may be run at very high resolution.

@@

We broadly commented on the hierarchical formation of galaxies. Over time, galaxies that merge with other galaxies should become well-mixed in the host halo. That is, there should be an underlying smooth distribution of mass in which all of the subhaloes reside. A key prediction of $\Lambda$CDM is that this smooth distribution should have a universal density profile in the limit that dark matter dominates the Universe \citep{nfw}. Furthermore, the dark matter should have a density distribution somewhat resembling a squashed football, a spheroid flattened on two axes, in general. There are, in fact, the special cases of being either perfectly spherical, prolate, or oblate. Whether or not this picture holds observationally in real galaxies is an open problem. The density profile proposed by \citet{nfw} is cuspy at the center, whereas some observations are more consistent with a flatter central density \citep{de_blok_core_cusp}. Astronomers have dubbed this the Core-Cusp Problem.

Besides the Core-Cusp Problem in the inner parts of dark matter haloes, one thing that is clear is that $\Lambda$CDM haloes should have many subhaloes. As the general theory for the formation of galaxies goes, these subhaloes were at one point independent perturbations that have since merged into the Milky Way's halo, and should have their own stellar and gas content. This means that we expect many massive subhaloes to be directly observable. Another triumph of simulations is putting an actual number on how many subhaloes we expect from $\Lambda$CDM. Surprisingly, far fewer subhaloes are actually observed \citep{mooresubhalos, Klypin1999, springel2008}. This discrepancy has since been dubbed the Missing Satellite Problem, and is an open problem for $\Lambda$CDM and astronomy.

@@

There is also a stellar component of the halo, which contains remnants of old mergers and star clusters known as globular clusters. This section is organized by talking about tidal disruption of subhaloes in the Galactic halo, the dynamics of the central part of the Galaxy, and vertical structure in the Galactic disc. How the cosmological environment effects these observables is the bulk of this thesis, and we motivate here why these aspects are the subject of focus.

@@

In the hierarchical formation of galaxies, smaller galaxies merge with larger galaxies and become a part of the larger galaxy's halo. As subhaloes become more a part of the host system, they experience tidal forces that stretch and squeeze their mass content. Dark matter and stellar material is stripped from these merging galaxies over time to form streams of tidal debris. In the long run, this debris will equlibrate with the stellar halo. It should be noted that the same effects apply to globular clusters too.

@@

The orbits of streams are not the only way to learn about the Milky Way's dark matter halo. It is worth mentioning that there is the possibility of using the fine structure of stellar streams to learn more about the subhalo population in the halo.  Gaps form when subhaloes pass through tidal streams, and this is one direction that is being explored for leveraging stellar stream data (e.g. \citet{erkal_2016_stream_gaps}).

@@

The Milky Way's massive subhaloes are most likely inducing non-planar density patterns in the disc, and we would like to understand the nature of these patterns through simulation. Unfortunately, we need the high resolution of isolated galaxy simulations to study these effects, and the realism of an \textit{ab initio} Galactic halo.  This has motivated a long history of attempts to bridge the gap between {\it ab initio} cosmological simulations and those of isolated galaxies.  A particularly simple class of simulations involves the interaction of a disc
with a single satellite galaxy or dark matter subhalo.  For example \citet{kazantzidis2008} perform simulations in which a thin Milky Way-like disc is subjected to a series of encounters with a satellite.  The masses and orbital parameters of the satellites are motivated by substructure found in a halo appropriate to a Milky Way-like galaxy from a cosmological simulation.  Their simulation demonstrated that satellite encounters lead to general disc heating as well as distinctive disc features such as bars, spiral structure, flares, and rings.  \citet{purcell2011} model the response of the Milky Way to the gravitational effects of
the Sgr dSph by simulating disc-dwarf encounter for different choices of the Sgr progenitor's mass.  Their conclusion is that Sgr may have triggered the development the spiral structure seen today in the Milky Way.  This approach has gained considerable traction in the last couple years, with authors using single-satellite encounters to explain the existence of low-latitude streams and other vertical structure in the Milky Way's disc \citep[for example]{widrow_2014, dlv_2015, donghia_2016, laporte_2016, laporte_2018}. One particularly interesting feature arising from these works is that stars may occasionally be kicked out of the galactic disc to form kicked-up disc populations \citep{johnston_kud_review, laporte_2019_feathers}. Such populations may explain low-latitude Milky Way structures, and possibly kinematic data of M31 \citep{dorman_2013_m31}.

@@

Of course, the disc of the Milky Way lives within a population of satellite galaxies and, quite possibly, pure dark matter subhaloes, and single-satellite encounters do not describe the complexity of the cosmological environment \citep{Klypin1999,mooresubhalos,springel2008}.  With this in mind \citet{Font2001} simulated the evolution of an isolated disc-bulge-halo model where the halo was populated by several hundred subhaloes.  They concluded that that substructure played only a minor role in heating the disc, a result that would seem in conflict with those of the \citet{kazantzidis2008} that would come later.  Numerical simulations by \citet{gauthier2006, dubinski2008} sheds some light on this discrepancy.  In these simulations, 10\% of the the halo mass in an isolated disc-bulge-halo system is replaced by subhaloes with a mass distribution motivated by the cosmological studies of \citet{gao2004}.  In the \citet{gauthier2006} simulation, little disc heating occurs during the first 5 Gyr, at which point satellite interactions provoke the formation of a strong bar, which in turn leads to significant heating.  Not surprisingly, when the experiment is repeated with different initial conditions for the satellites the timing of bar formation can vary.  These experiments also suggest that a large fraction of the mass initially in subhaloes is tidally stripped leaving behind a complex system of tidal debris.

The aforementioned simulations, although more realistic than single-satellite encounters, have three main problems.  First, they don't allow for halo triaxiality.  Secondly, the disc is initialized at its present-day mass whereas real stellar discs form over several Gyr.  Finally, the subhalo populations are inserted into the halo by hand.  While the mass and spatial distributions of the subhaloes are motivated by cosmological simulations, they may not capture important properties of realistic haloes. This has historically been solved by modifying a traditional \textit{ab initio} simulation. The zoom-in technique (a Monte Carlo adaptive mesh) is used to study the evolution of a disc in a single halo at high resolution.  This is accomplished by resampling the initial conditions at a sequence of higher resolutions after identifying an area of interest. With sufficiently realistic feedback, impressively realistic results can be obtained. For instance, \citet{gomez_2017} presented a study of vertical structure in such simulations. In these simulations, the authors were able to look at several Milky Way-like galaxies in a variety of environments at resolutions that rival isolated galaxy simulations. These realistic studies have opened the door to a wide variety of questions pertaining to the Milky Way's interaction with the cosmological environment.

@@

The general idea for growing a disc is to divide a zoom-in simulation into three phases.  During the first, a cosmological simulation of a region of interest is run, and a halo is identified for study.  A rigid disc potential is then slowly grown in this halo, allowing the halo particles to respond adiabatically to its time-varying potential.  Finally, the rigid disc is replaced by live particles and the final phase of the simulation proceeds with live disc and halo particles.  \citet{debuhr_2012} use this scheme to introduce stellar discs into dark matter haloes from the Aquarius Project \citep{springel2008}.  They add a rigid disc potential at redshift $z=1.3$ with a mass parameter for the disc that grows linearly with the scale factor until $z=1$ when it reaches its final value.  The disc potential is initially centered on the potential minimum of the halo and oriented so that its symmetry axis points along either the minor or major axis of the halo.  During the rigid disc phase, the centre of the potential moves due to its interaction with the halo (Newton's 3rd law) so that linear momentum is conserved.  However, the orientation of the disc is held fixed and therefore angular momentum isn't.  To initialize the live disc, \citet{debuhr_2012} approximate the halo potential as a flattened logarithmic potential and then determine the disc distribution function by solving the Jeans equations in axisymmetry.

@@

The common theme between these topics as follows. When studied in isolated galaxy simulations, stellar discs display evolution according to their ``nature". We learn about what properties intrinsic to the disc change its evolution. In a fully cosmological environment, external factors will change the natural evolution of the stellar system. We refer to this aspect as  ``nurture". Real stellar systems evolve as a combination of the two, and we seek to understand how each plays a role in realistic environments.

As such, the thesis is organized in the following manner. In Chapter~\ref{ch:background}, we reduce the necessary theoretical background to understand this work to a succinct summary. Following this, in Chapter~\ref{ch:paper_i}, we explore the first peer-reviewed paper on which this thesis is based. It outlines the disc insertion scheme and comments on how the dynamic nature of the Milky Way's stellar disc should affect its inner halo. Chapter~\ref{ch:paper_ii} presents the second published paper which focuses on the question of how stable disks are against forming bars in a cosmological setting. A final paper is presented in Chapter~\ref{ch:paper_iii}, where we look at how disks form vertical structure in a cosmological setting. We conclude with summaries and ideas for future work in Chapter~\ref{ch:conclusion}.

### 2019-09-23 Thesis/Paper_I.tex
Final commit

In Fig.\,\ref{fig:axis_ratios} we show the minor-to-major ($c_r/a_r$)
and intermediate-to-major ($b_r/a_r$) axis ratios as a function of
radius for both the $z=1$ and $z=0$ snapshots.  The axis ratios are
calculated by diagonalizing the moment of inertia tensor in
linearly-spaced radial shells.  The DMO halo is triaxial with
$c_r/a_r\simeq 0.75$ and $b_r/a_r\simeq 0.85$.  Note that the axis
ratio profiles are smoother at $z=0$ than at $z=1$, which supports the
observation that the halo has settled into a more relaxed state over
the past $7$ or so billion years.  In general, discs tend to make
halos more spherical, a result that has been known for some time from
both collisionless and hydrodynamical simulations
\citep[e.g.][]{dubinski1994_ApJ431_617,Zemp2012}.

>>>>

In Fig.\,\ref{fig:axis_ratios} we show the minor-to-major ($c_r/a_r$)
and intermediate-to-major ($b_r/a_r$) axis ratios as a function of
radius for both the $z=1$ and $z=0$ snapshots.  The axis ratios are
calculated by diagonalizing the moment of inertia tensor in
linearly-spaced radial shells.  The DMO halo is triaxial with
$c_r/a_r\simeq 0.75$ and $b_r/a_r\simeq 0.85$.  Note that the axis
ratio profiles are smoother at $z=0$ than at $z=1$, which supports the
observation that the halo has settled into a more relaxed state over
the past $7$ or so billion years.  In general, discs tend to make
haloes more spherical, a result that has been known for some time from
both collisionless and hydrodynamical simulations
\citep[e.g.][]{dubinski1994_ApJ431_617,Zemp2012}.
